%\section{Related Works}\label{section:related}

%\paragraph{{\bf Related Work}}

In the past, some efforts have been undertaken to enable \abft support in MPI. 
FT-MPI~\cite{fagg2000ft} was an MPI-1 implementation which proposed
changes to the MPI semantics to enable repairing communicators, thus
re-enabling communications for applications damaged by failures. This
approach has proven successful and applications have been
implemented using FT-MPI. However, these modifications were not adopted by
the MPI standardization body, and the resulting lack of portability
undermined user adoption for this fault tolerant solution.

%  One of
% the solutions implemented in FT-MPI was to introduce a new MPI\_Errhandler
% called MPI\_ERRORS\_BLANK. This MPI\_Errhandler replaced the position of the
% failed processes inside a communicator with MPI\_PROC\_NULL. By using this
% semantic, the remaining MPI calls could function normally as communication with
% MPI\_PROC\_NULL always succeeds. When FT-MPI encountered a fault, it destroyed
% all MPI Communicators and required that the application recreate them to account
% for the failed processes. While this was a useful step in allowing the most
% level of flexibility from the application's perspective, it made recovery very
% complex and added a large overhead.

% FT-MPI is another form of fault tolerance that can be successful for some, but in
% applications that require that all processes be running to reach successful
% completion, having a hole in the communicator is not a valid solution. The
% failed processes need to somehow be recovered.

%\TBD{Anything needed from the QR side of things?}

During the process that recently resulted in the MPI-3 Standard, a specific
working group was assembled to investigate the issues of Fault Tolerance support
in MPI. Some of the early results are outlined in the following
publication~\cite{Hursey11MPI3FT}. Late in the process, promising results had
been demonstrated toward effective support of process failures and continued MPI
operations with acceptable overhead~\cite{DBLP:conf/pvm/BlandBHHBD12}. However,
these propositions were in too early a state to meet the calendar requirements
of MPI-3 and their adoption (and according availability in production systems)
is, at best, postponed to the restart of the MPI Forum toward the next version
of the MPI standard. 

In %the article by W. Gropp and E. Lusk in 2004
\cite{Gropp:2004:FTM:1080704.1080714}, the authors discuss alternative
or slightly modified interpretations of the MPI Standard that enable
some forms of fault tolerance. 
One core idea is that process failures happening in another MPI world,
connected only through an inter-communicator, should not prevent the
continuation of normal operations. The complexity of this approach, for
both the implementation and users, has prevented these ideas from having
a practical impact.

In the \cof approach, the only requirement from the MPI implementation
is that it does not forcibly kill the living processes without returning
control. No strong support from the MPI stack is required, in particular
the state of the library is left undefined and further communication is
impossible without a restart. This simplicity has enabled us to actually
implement our proposition, and then experimentally support and evaluate
a real \abft application. Similarly, little effort would be required to
extend MPICH-2 to support \cof (see Section 7 of the MPICH
Readme\footnote{\url{http://www.mpich.org/documentation/guides/files/mpich2-1.5-README.txt}}).




