\section{Conclusion}
\label{sect:conclusion}

Many responsible voices agree that sharp increases in the volatility of future,
extreme scale computing platforms are likely to jeopardize our ability to use
them to maximize the research productivity of advanced long-running applications
that deliver meaningful scientific results. Moreover, it become clear that any
lightweight techniques developed to address this volatility must be supported in the
programming and execution model. Since MPI is currently, and will likely
continue to be -- in the medium-term -- both the de-facto programming model for
distributed applications and the default execution model for large scale
platforms running at the bleeding edge, MPI is the place in the software
infrastructure where semantic and run-time support for application faults needs
to be provided.

The \ulfm proposal is a careful but important step forward toward accomplishing
this goal.  It not only delivers support for a number of new and innovative
resilience techniques, it provides this support through a simple,
straightforward and familiar API that requires minimal modifications of the
underlying MPI implementation. It maintains the backward compatibility with
previous versions of the MPI standard, so that non fault-tolerant applications
(legacy or otherwise) are supported without any changes to the code. In same
time it provided a low-level portable API, that will allow other fault tolerant
approaches to be implemented as libraries. Perhaps most significantly,
applications can use \ulfm-enabled MPI without experiencing any degradation in
their performance, as we demonstrate in this paper.

Several applications, ranging from Master-Worker to tightly coupled, are
currently being refactored to take advantage of the semantics in this
proposal. Beyond applications, the expressivity of this proposal is 
investigated in the context of providing extended fault tolerance 
models as convenience, portable libraries.
